---
title: "Taxonomy Assignment"
Author: "Sarah Lucas"
---
# Setup
## Setup environment
```{r}
.cran_packages <- c("tidyverse" )
.bioc_packages <- c("dada2")

.inst <- .cran_packages %in% installed.packages()
if(any(!.inst)) {
   install.packages(.cran_packages[!.inst])
}
.inst <- .bioc_packages %in% installed.packages()
if(any(!.inst)) {
   BiocManager::install(.bioc_packages[!.inst], quietly = FALSE)
}
# Load packages into session, and print package version
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
```

```{r}
seqtab <- readRDS(file = "16S_output/seqtab.rds")
```

# Assign Taxonomy 
## SILVA train set
Download the latest SILVA training sets here: https://benjjneb.github.io/dada2/training.html
HOMD files are saved in the Training Set folder
```{r}
# Assign taxonomy SILVA Train Set
# Change path of training set to wherever fasta file is located
tax_silva <- assignTaxonomy(seqtab, "~/Documents/MICaB/Hunter_Lab/taxonomyTrainingSets/silva_nr_v132_train_set.fa.gz", multithread=TRUE)
colnames(tax_silva) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")
```

## Assign species from the SILVA dataset
Here, allowMultiple is set to FALSE, meaning that the sequence must only have 1 exact match to be assigned to species level.
```{r}
## Add species assignment to taxonomy table: https://benjjneb.github.io/dada2/assign.html#species-assignment
# Change path of training set to wherever fasta file is located
tax_species_silva <- addSpecies(tax_silva, "~/Documents/MICaB/Hunter_Lab/taxonomyTrainingSets/silva_species_assignment_v132.fa.gz", 
                                verbose=TRUE, 
                                allowMultiple = FALSE)
colnames(tax_species_silva)  <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
unname(head(tax_species_silva))
```

## Assign Species with HOMD train set
Here, allowMultiple is set to FALSE, meaning that the sequence must only have 1 exact match to be assigned to species level.
```{r}
## Add species assignment to taxonomy table: https://benjjneb.github.io/dada2/assign.html#species-assignment
tax_species_silva_HOMD <- addSpecies(tax_silva, "~/Documents/MICaB/Hunter_Lab/taxonomyTrainingSets/HOMD_16S_rRNA_RefSeq_V15.1.p9_dada2_addspecies.fasta", 
                                     verbose=TRUE, 
                                     allowMultiple = FALSE)
colnames(tax_species_silva_HOMD) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species_HOMD")
unname(head(tax_species_silva_HOMD))
```

## Combine both species assignments
I want to keep the SILVA assigned species, but then add the HOMD assigned species as well.
```{r}
#make rownames for ASVs for each species level taxa table
tax_species_silva_df <- as.data.frame(tax_species_silva) %>%
  rownames_to_column('ASV')
tax_species_silva_df[tax_species_silva_df == "cf."] <- NA
tax_species_silva_df[tax_species_silva_df == "sp."] <- NA
dim(tax_species_silva_df)

tax_species_silva_HOMD_df <- as.data.frame(tax_species_silva_HOMD) %>%
  rownames_to_column('ASV')
tax_species_silva_HOMD_df[tax_species_silva_HOMD_df == "cf."] <- NA
tax_species_silva_HOMD_df[tax_species_silva_HOMD_df == "sp."] <- NA
dim(tax_species_silva_HOMD_df)

# Use a full join to keep all rows and all columns x and y; NA for not matching values.
# by is set to ASV, Kingdom, Phylum, Class, Order, Family, Genus, so the join must match on all these levels.
tax_species_silva_HOMD_df_join <- full_join(x = tax_species_silva_df, 
                                            y = tax_species_silva_HOMD_df, 
                                            by = c("ASV","Kingdom", "Phylum", "Class", "Order", "Family", "Genus"))
dim(tax_species_silva_HOMD_df_join)
```
# Add a column to distinguish whether the species assignment came from SILVA, HOMD, or both
```{r}
tax_species_silva_HOMD_df_join <- tax_species_silva_HOMD_df_join %>%
  mutate(DB = ifelse(!is.na(Species) & !is.na(Species_HOMD), "Both",
                     ifelse(!is.na(Species), "SILVA",
                            ifelse(!is.na(Species_HOMD), "HOMD", NA))))
```

```{r}
HOMD <- filter(tax_species_silva_HOMD_df_join, DB == "HOMD")
HOMD
```

```{r}
# If there is no species assignment in the Species_SILVA column, check the Species_HOMD column. If the Species_HOMD column says NA or sp. assign NA.
tax_species_silva_HOMD_df_join_newspecies <- tax_species_silva_HOMD_df_join %>%
  mutate(SpeciesCombo = ifelse(!is.na(Species), as.character(Species), as.character(Species_HOMD))) %>%
  select(ASV, Kingdom, Phylum, Class, Order, Family, Genus, SpeciesCombo, DB) %>%
  dplyr::rename(Species = SpeciesCombo) %>%
  column_to_rownames('ASV') %>%
  as.matrix()
```
```{r}
# Write to disk
saveRDS(tax_species_silva_HOMD_df_join_newspecies, "16S_output/tax_species_final_silva_HOMD.rds")
```

# Evaluate Tax Addition using HOMD

Make a table where the SILVA species assignment was NA, but the HOMD assignment was not. Count how many rows (ASVs).
```{r}
speciesCompare <- tax_species_silva_HOMD_df_join %>%
  filter(is.na(Species) & !is.na(Species_HOMD))
dim(speciesCompare)
```

How many rows were not NA in the Species column for the SILVA species dataframe?
```{r}
length(which(!is.na(tax_species_silva_df$Species)))
```
How many rows were not NA in the Species column for the HOMD species dataframe?
```{r}
length(which(!is.na(tax_species_silva_HOMD_df$Species_HOMD)))
```
Do these numbers remain consistent when the tables are joined together?
```{r}
length(which(!is.na(tax_species_silva_HOMD_df_join$Species)))
length(which(!is.na(tax_species_silva_HOMD_df_join$Species_HOMD)))
```
Yes, they do.