---
title: "Abundance Analysis - Top Taxa"
output: word_document
---
# Setup
```{r}
.cran_packages <- c("tidyverse","gridExtra", "devtools", 
                    "adaptiveGPCA", "ade4", "vegan", 
                    "devtools", "ggthemes", "naniar", 
                    "scales", "extrafont", "data.table",
                    "ggpubr", "cowplot")
.bioc_packages <- c("phyloseq", "BiocStyle","Biobase")

# Install CRAN packages (if not already installed)
.inst <- .cran_packages %in% installed.packages()
if (any(!.inst)){
  install.packages(.cran_packages[!.inst],repos = "http://cran.rstudio.com/")
}

.inst <- .bioc_packages %in% installed.packages()
if (any(!.inst)){
   BiocManager::install(.bioc_packages[!.inst], quietly = FALSE)
}

# Load packages into session, and print package version
sapply(c(.cran_packages, .bioc_packages), require, character.only = TRUE)
sapply(c(.cran_packages, .bioc_packages), package.version)

library(ampvis2)
```


# Load the data:
```{r}
# Rarified filtered data used in ordination and picrust
data <- readRDS("16S_output/phyloseq_unsupervised_filtered_rare.rds")
data
```
Move first column of sample data to be SAMPLE_NAME
```{r}
samdf <- data.frame(sample_data(data)) %>%
  select(SAMPLE_NAME, everything())
sample_data(data) <- samdf
```

# Description of the data

## Phyla level composition of samples
```{r}
CRS <- subset_samples(data, DIAG_CRS == "CRS") %>% filter_taxa(.,function(x) sum(x) > 0, TRUE)
CRS
nonCRS <- subset_samples(data, DIAG_CRS == "Non-CRS") %>% filter_taxa(.,function(x) sum(x) > 0, TRUE)
nonCRS
```
What is the relative abundance of each phylum?
```{r}
CRS.phy <- tax_glom(CRS, taxrank = "Phylum")
tax.count.CRS <-
  data.frame(taxa_sums(CRS.phy), tax_table(CRS.phy)[, 2])
rownames(tax.count.CRS) = NULL
colnames(tax.count.CRS)[1] <- c("Abundance")
tax.count.CRS$Percent <-
  round(tax.count.CRS$Abundance / sum(tax.count.CRS$Abundance) * 100, 4)
library(plyr)
Phylum_df <-
  tax.count.CRS[with(tax.count.CRS, order(-Percent)),]

#how much do the top 5 phyla contribute to total abundance?
top5PhyCRS <- Phylum_df[1:5,]
round(sum(top5PhyCRS$Percent), 2)
```

```{r}
top5PhyCRS
```
```{r}
nonCRS.phy <- tax_glom(nonCRS, taxrank = "Phylum")
tax.count.nonCRS <-
  data.frame(taxa_sums(nonCRS.phy), tax_table(nonCRS.phy)[, 2])
rownames(tax.count.nonCRS) = NULL
colnames(tax.count.nonCRS)[1] <- c("Abundance")
tax.count.nonCRS$Percent <-
  round(tax.count.nonCRS$Abundance / sum(tax.count.nonCRS$Abundance) * 100,
        4)
library(plyr)
Phylum_df <-
  tax.count.nonCRS[with(tax.count.nonCRS, order(-Percent)),]

#how much do the top 5 phyla contribute to total abundance?
top5PhyNonCRS <- Phylum_df[1:5,]
round(sum(top5PhyNonCRS$Percent), 2)
```
```{r}
top5PhyNonCRS
```

## Description of data: genus-level summary of CRS samples
```{r}
#What is the number of Genera found?
length(get_taxa_unique(CRS, taxonomic.rank="Genus"))
```
```{r}
#what are the abundance levels of each genus?
CRS.genus <- tax_glom(CRS, taxrank = "Genus")
tax.count.CRS <-
  data.frame(tax_table(CRS.genus)[, 2:6], taxa_sums(CRS.genus))
rownames(tax.count.CRS) = NULL
colnames(tax.count.CRS) <-
  c("Phylum", "Class", "Order", "Family", "Genus", "Abundance")
tax.count.CRS$Percent <-
  round(tax.count.CRS$Abundance / sum(tax.count.CRS$Abundance) * 100, 4)
Genus_df_CRS <-
  tax.count.CRS[with(tax.count.CRS, order(-Percent)),]

#how much do the top 10 genera contribute to total abundance?
top10GeneraCRS <- Genus_df_CRS[1:10,]
round(sum(top10GeneraCRS$Percent), 3)
```
78.2% of the seqeunces can be attributed to the top 10 genera in the CRS samples. What are they?
```{r}
###How diverse are the top 10 genera? i.e., how many species are there per genus?
top10GeneraCRS <- as.vector(Genus_df_CRS$Genus[1:10])
Diversity.list.CRS <- vector("list", 10)
names(Diversity.list.CRS) <- top10GeneraCRS

for (i in 1:length(top10GeneraCRS)) {
  physub = subset_taxa(CRS, Genus == top10GeneraCRS[i])
  physub = prune_taxa(taxa_sums(physub) > 0, physub)
  Diversity.list.CRS[[i]] <- physub
}

#compute the number of taxa in each element of the list
NtaxaCRS <- data.frame(unlist(lapply(Diversity.list.CRS, ntaxa)))

colnames(NtaxaCRS) <- "N.Species"
#Make a table with percent abundance and number of taxa
genus.tab.CRS <- data.frame(Genus_df_CRS[1:10, ], NtaxaCRS)
genus.tab.CRS
```

## Description of data: genus-level summary of non-CRS samples
```{r}
#What is the number of Genera found?
length(get_taxa_unique(nonCRS, taxonomic.rank="Genus"))
```
There are 153 unique genera in the nonCRS sample dataset
```{r}
#what are the abundance levels of each genus?
nonCRS.genus <- tax_glom(nonCRS, taxrank = "Genus")
tax.count.nonCRS <-
  data.frame(tax_table(nonCRS.genus)[, 2:6], taxa_sums(nonCRS.genus))
rownames(tax.count.nonCRS) = NULL
colnames(tax.count.nonCRS) <-
  c("Phylum", "Class", "Order", "Family", "Genus",  "Abundance")
tax.count.nonCRS$Percent <-
  round(tax.count.nonCRS$Abundance / sum(tax.count.nonCRS$Abundance) * 100,
        4)
library(plyr)
Genus_df_nonCRS <-
  tax.count.nonCRS[with(tax.count.nonCRS, order(-Percent)),]

#how much do the top 10 genera contribute to total abundance?
top10GeneraNonCRS <- Genus_df_nonCRS[1:10,]
round(sum(top10GeneraNonCRS$Percent), 3)
```
84.5% of the seqeunces can be attributed to the top 10 genera in the CRS samples. What are they?
```{r}
###How diverse are the top 10 genera? i.e., how many species are there per genus?
top10GeneraNonCRS <- as.vector(Genus_df_nonCRS$Genus[1:10])
Diversity.list.nonCRS <- vector("list", 10)
names(Diversity.list.nonCRS) <- top10GeneraNonCRS

for (i in 1:length(top10GeneraNonCRS)) {
  physub = subset_taxa(nonCRS, Genus == top10GeneraNonCRS[i])
  physub = prune_taxa(taxa_sums(physub) > 0, physub)
  Diversity.list.nonCRS[[i]] <- physub
}

#compute the number of taxa in each element of the list
NtaxaNonCRS <-
  data.frame(unlist(lapply(Diversity.list.nonCRS, ntaxa)))

colnames(NtaxaNonCRS) <- "N.Species"
#Make a table with percent abundance and number of taxa
genus.tab.nonCRS <- data.frame(Genus_df_nonCRS[1:10, ], NtaxaNonCRS)
genus.tab.nonCRS
```
# Phylum level abundance
## Stacked Bargraphs
```{r}
dataPhylumPropMelt <- data %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  psmelt()
# Make anything with an abundance less than 1% "Other"
dataPhylumPropMelt$Phylum <- as.character(dataPhylumPropMelt$Phylum)
dataPhylumPropMelt <- mutate(dataPhylumPropMelt, PhylumOther = ifelse(Abundance < 1, "Other < 1%", Phylum))

bugcolors <- read_csv("CRS_Taxa_Colors.csv")
dataPhylumPropMelt <- left_join(dataPhylumPropMelt, bugcolors, by = c("PhylumOther" = "Taxon")) %>%
  select(SAMPLE_NAME, DIAG_CRS, SINUS_FACTORS_NUMFESS, SINUS_FACTORS_PRIORFESS, contains("ABX"), Phylum, PhylumOther, Abundance, Color)
#Separate CRS and non CRS datasets, arrange, and get sample names for plotting
dataPhylumPropMeltCRS <- dataPhylumPropMelt %>%
  filter(DIAG_CRS == "CRS") %>%
  dplyr::arrange(DIAG_CRS, PhylumOther, desc(Abundance)) 

dataPhylumPropMeltNonCRS <- dataPhylumPropMelt %>%
  filter(DIAG_CRS == "Non-CRS") %>%
  dplyr::arrange(DIAG_CRS, PhylumOther, desc(Abundance)) 

  
orderXCRS <- unique(dataPhylumPropMeltCRS$SAMPLE_NAME) 
labelsXCRS <- unique(dataPhylumPropMeltCRS$SAMPLE_NAME)
orderXNonCRS <- unique(dataPhylumPropMeltNonCRS$SAMPLE_NAME) 
labelsXNonCRS <- unique(dataPhylumPropMeltNonCRS$SAMPLE_NAME)

```
```{r}
colorTax <- unique(dataPhylumPropMeltCRS$PhylumOther)
colorHex <- unique(dataPhylumPropMeltCRS$Color)

barplot.phylum.prop.CRS <- ggplot(data=dataPhylumPropMeltCRS, aes(x = SAMPLE_NAME, y = Abundance, fill = PhylumOther)) +
  geom_bar(aes(), stat="identity", position="stack") + 
  scale_x_discrete(limits = orderXCRS,
                   labels = labelsXCRS) +
  theme_minimal(base_size = 14) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        plot.margin = margin(0,0,0,0, "cm"),
        panel.grid = element_blank()) +
  scale_fill_manual(breaks = colorTax, values = colorHex)
barplot.phylum.prop.CRS
```

```{r}
colorTaxNonCRS <- unique(dataPhylumPropMeltNonCRS$PhylumOther)
colorHexNonCRS <- unique(dataPhylumPropMeltNonCRS$Color)

barplot.phylum.prop.NonCRS <- ggplot(data=dataPhylumPropMeltNonCRS, aes(x = SAMPLE_NAME, y = Abundance, fill = PhylumOther)) +
  geom_bar(aes(), stat="identity", position="stack") + 
  scale_x_discrete(limits = orderXNonCRS,
                   labels = labelsXNonCRS) +
  theme_minimal(base_size = 14) + 
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_text(color = "black"),
        plot.margin = margin(0,0,0,0, "cm"),
        panel.grid = element_blank()) +
  scale_fill_manual(breaks = colorTaxNonCRS, values = colorHexNonCRS)
barplot.phylum.prop.NonCRS
```
```{r, fig.width=25, fig.height=8}
Phylum.Prop.Legend <- get_legend(barplot.phylum.prop.CRS)
Phylum.Prop.Grid <- plot_grid(barplot.phylum.prop.CRS + theme(legend.position = ""),
                              barplot.phylum.prop.NonCRS + theme(legend.position = ""), 
                              Phylum.Prop.Legend, nrow = 1, rel_widths = c(4, 1.5, 1))
Phylum.Prop.Grid
```
```{r}
#ggsave(Phylum.Prop.Grid,
#       filename = "16S_output/figures/PhylumPropBarGrid.pdf",
#       device = "pdf",
#      height = 8,
#       width = 25)
```

## Boxplots
```{r}
Phylum.prop <- data %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  psmelt() %>%
  filter(Abundance > 0)
```
```{r fig.height=10, fig.width=16}
p <- ggboxplot(Phylum.prop, 
              x = "DIAG_CRS", y = "Abundance",
              add = "jitter",
              facet.by = "Phylum",
              scales = "free"
              ) + 
  stat_compare_means(method = "wilcox.test") + 
  scale_y_log10()
p
```


# Genus level abundance analysis
Convert phyloseq to ampvis2 object
```{r}
#Combine OTU abundance table and taxonomy table from the phyloseq object "my_phyloseq_object":
obj <- data
# Make short names for OTUs - facilitates later plotting ease when amp_heatmap has the option tax_empty = "best" (for some reason)
taxa_names(obj) <- paste0("ASV", seq(ntaxa(obj)))
# Fix OTU table layout for exporting. taxa_as_rows = FALSE was not working.
Totu_table =t(otu_table(obj))
otu_table(obj)=Totu_table
#export OTU table from phyloseq object
otutable <- data.frame(OTU = rownames(phyloseq::otu_table(obj)@.Data),
                       phyloseq::otu_table(obj)@.Data,
                       phyloseq::tax_table(obj)@.Data,
                       check.names = FALSE
                       )
otutable <- otutable %>% select(-OTU, 
                                -TaxName, # Remove special taxa formatting
                                -Genus_Species2, # Remove special taxa formatting
                                -DB)

#Extract metadata from the phyloseq object:
metadata <- data.frame(phyloseq::sample_data(obj), 
                       check.names = FALSE
                       )
metadata <- rownames_to_column(metadata, var = "SAMPLE_ID")

# Extract phylogenetic tree from phyloseq object:
#phytree <- phyloseq::phy_tree(obj)

#Load the data with amp_load:
ampvis.obj <- amp_load(otutable, metadata#,
                       #tree = phytree
                        )
ampvis.obj

# Split inot two datasets for CRS and Non-CRS samples
ampvis.CRS <- amp_subset_samples(ampvis.obj, DIAG_CRS == "CRS")
ampvis.CRS
ampvis.NonCRS <- amp_subset_samples(ampvis.obj, DIAG_CRS == "Non-CRS")
ampvis.NonCRS
```
## Heatmaps
### CRS
```{r}
heatmap.CRS <- amp_heatmap(ampvis.CRS,
                           tax_aggregate = "Genus",
                           #tax_add = "OTU",
                           #tax_empty = "Family",
                           tax_show = 15,
                           plot_values = FALSE,
                           plot_colorscale = "log10",
                           normalise = TRUE,
                           order_x_by = "cluster",
                           color_vector = c("whitesmoke", "black"), #"#347A88"
                           ) + 
  theme_minimal(base_size = 14) +
  theme(#axis.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust=1, vjust=.5),
        axis.text.y = element_text(face = "italic", color = "black"),
        axis.ticks = element_blank(),
        panel.grid = element_blank(),
        legend.position = "",
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm")
  ) + scale_x_discrete(expand = c(0,3)) + scale_y_discrete(expand = c(0,0)) + coord_fixed(ratio=2/1)
heatmap.CRS
```
```{r}
#ggsave(heatmap.CRS, 
#       filename = "16S_output/figures/taxaHeatPlotCRS.pdf",
#       device = "pdf",
#       height = 3)
```
Get abundance dataframe for the heatmap
```{r}
heatmap.CRS.df <- amp_heatmap(ampvis.CRS,
                           textmap = TRUE,
                           tax_show = 15,
                           tax_aggregate = "Genus",
                           #tax_add = "OTU",
                           plot_values = FALSE,
                           normalise = TRUE,
                           order_x_by = "cluster",
                           )
heatmap.CRS.df
```
Make a dataframe with just the sample names and numbers in the heatmap order for plotting culture data:
```{r}
CRS.heat.names <- colnames(heatmap.CRS.df)
heatmap.CRS.df.names <- tibble("subject_id"=CRS.heat.names)
#write.csv(heatmap.CRS.df.names, "../data/16S/CRS_heatmap_names_ordered.csv")
```


Make barplot from averages of these numbers
```{r fig.height=3, fig.width=1.5}
heatmap.CRS.df.meanRA <- heatmap.CRS.df %>%
  rownames_to_column("Genus") %>%
  pivot_longer(-Genus, names_to = "Sample", values_to = "RA") %>%
  group_by(Genus) %>%
  dplyr::summarise(meanRA = mean(RA)) %>%
  dplyr::arrange(desc(meanRA))
# Make a barplot
meanRAbarCRS <- ggplot(heatmap.CRS.df.meanRA, aes(x = reorder(Genus,meanRA), y = meanRA)) +
  geom_col(fill = "lightgrey", color = "white", width = 1) +
  coord_flip() +
  theme_minimal(base_size = 10) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(color = "black"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank())
meanRAbarCRS
```
```{r}
#ggsave(meanRAbarCRS, filename = "16S_output/figures/meanRAbarplotCRS.pdf",
#       device ="pdf",
#       height = 3.1,
#      width = .8)
```


### Non-CRS
```{r}
heatmap.NonCRS <- amp_heatmap(ampvis.NonCRS,
                           tax_aggregate = "Genus",
                           tax_show = 15,
                           plot_values = FALSE,
                           plot_colorscale = "log10",
                           normalise = TRUE,
                           order_x_by = "cluster",
                           color_vector = c("whitesmoke", "black") #"#347A88"
                           ) + 
  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(angle=90),
        #axis.text.x = element_blank(),
        axis.text.y = element_text(face="italic", color = "black"),
        axis.ticks = element_blank(),
        legend.position = "",
        panel.grid = element_blank(),
        plot.margin = margin(0,0,0,0, unit = "cm"),
        ) + 
  scale_x_discrete(expand = c(0,3)) + 
  scale_y_discrete(expand = c(0,0)) + 
  coord_fixed(ratio = 2/1)


heatmap.NonCRS
```
```{r}
heatmap.NonCRS.df <- amp_heatmap(ampvis.NonCRS,
                           textmap = TRUE,
                           tax_show = 15,
                           tax_aggregate = "Genus",
                           plot_values = FALSE,
                           normalise = TRUE,
                           order_x_by = "cluster",
                           ) 
heatmap.NonCRS.df
```
```{r}
#ggsave(heatmap.NonCRS, 
#       filename = "16S_output/figures/taxaHeatPlotNonCRS.pdf",
#       device = "pdf",
#       height = 3.0)
```

Make barplot from averages of these numbers
```{r}
heatmap.NonCRS.df.meanRA <- heatmap.NonCRS.df %>%
  rownames_to_column("Genus") %>%
  pivot_longer(-Genus, names_to = "Sample", values_to = "RA") %>%
  group_by(Genus) %>%
  dplyr::summarise(meanRA = mean(RA)) %>%
  dplyr::arrange(desc(meanRA))
# Make a barplot
meanRAbarNonCRS <- ggplot(heatmap.NonCRS.df.meanRA, aes(x = reorder(Genus,meanRA), y = meanRA)) +
  geom_col(fill = "lightgrey", color = "white", width = 1) +
  coord_flip() +
  theme_minimal(base_size = 10) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_text(color = "black"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank())
meanRAbarNonCRS
```
```{r}
#ggsave(meanRAbarNonCRS, filename = "16S_output/figures/meanRAbarplotNonCRS.pdf",
#       device ="pdf",
#       height = 3.1,
#       width = 0.8)
```

### Plot Grid

```{r fig.height=3, fig.width=21}
heatmap_grid <- plot_grid(heatmap.CRS, meanRAbarCRS, heatmap.NonCRS, meanRAbarNonCRS, 
                          nrow = 1, 
                          align = "hv",
                          rel_widths = c(3,1,3,1))
heatmap_grid
```





### Whole dataset: CRS/Non-CRS
```{r}
heatmap.obj <- amp_heatmap(ampvis.obj,
                           group_by = "DIAG_CRS",
                           facet_by = "DIAG_CRS",
                           tax_aggregate = "Genus",
                           plot_values = TRUE,
                           #tax_add = "Phylum",
                           tax_show = 30,
                           plot_colorscale = "log10",
                           normalise = TRUE,
                           #order_x_by = "hclust",
                           color_vector = c("whitesmoke", "pink"),
                           ) + 
  theme_minimal(base_size = 10) + 
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        legend.position = "right")
heatmap.obj
```

# Table of prevalence and abundance
Get the prevalence, sequence abundance and mean relative abundance of each unique taxname for the top 20 taxa
## Names of top 20 taxa by relative abundance
```{r}
top20genera.df <- amp_heatmap(ampvis.obj,
                           textmap = TRUE,
                          tax_aggregate = "Genus",
                           tax_show = 20,
                           plot_values = FALSE,
                           normalise = TRUE
                           ) 
top20generaNames <- rownames(top20genera.df)
top20generaNames
```

## Agglomerate CRS and Non-CRS datasets by the "TaxName" level.
```{r}
CRS.taxname <- tax_glom(CRS, "TaxName")
CRS.taxname
nonCRS.taxname <- tax_glom(nonCRS, "TaxName")
nonCRS.taxname
```

## Calculate relative abundances for each taxname
```{r}
CRS.taxname.prop <- transform_sample_counts(CRS.taxname, function(x) x/sum(x) * 100)
nonCRS.taxname.prop <- transform_sample_counts(nonCRS.taxname, function(x) x/sum(x) * 100)
```

## Summarize relative abundances by the mean
Get a table of average relative abundances for each taxname, take the average across the samples, including zeros in the calculation 
(So data matches up with the ampvis heatmaps, which calculate average relative abundances including zeros)
```{r}
CRS.taxname.prop.melt <- psmelt(CRS.taxname.prop) %>%
  #filter(Abundance > 0) #%>%
  group_by(OTU, Phylum, Class, Order, Family, Genus, Species, TaxName) %>%
  dplyr::summarize(avgRelAbun = mean(Abundance)) %>%
  ungroup() %>%
  select(TaxName, avgRelAbun)
dim(CRS.taxname.prop.melt)
```
```{r}
nonCRS.taxname.prop.melt <- psmelt(nonCRS.taxname.prop) %>%
  #filter(Abundance > 0) %>%
  group_by(OTU, Phylum, Class, Order, Family, Genus, Species, TaxName) %>%
  dplyr::summarize(avgRelAbun = mean(Abundance)) %>%
  ungroup() %>%
  select(TaxName, avgRelAbun)
dim(nonCRS.taxname.prop.melt)
```

## Prevalence, feature counts, and percent of total sequences for each TaxName
Calculated from the original, non-transformed dataset. Bind to dataframe of mean relative abundances.

### CRS
```{r}
# Compute prevalence of each feature, store as data.frame
prevdf.taxname.CRS = apply(X = otu_table(CRS.taxname),
               MARGIN = ifelse(taxa_are_rows(CRS.taxname), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
# Add taxonomy and total read counts to this data.frame
prevdf.taxname.CRS = data.frame(Prevalence = prevdf.taxname.CRS,
                    Abundance = taxa_sums(CRS.taxname),
                    phyloseq::tax_table(CRS.taxname)@.Data) %>%
  mutate(PercentAbundance = round(Abundance/sum(Abundance) * 100, 4)) %>%
  rownames_to_column("ASV") %>%
  left_join(CRS.taxname.prop.melt, by = "TaxName") %>%
  column_to_rownames("ASV") %>%
  select(Prevalence, Abundance, PercentAbundance, avgRelAbun, TaxName, everything(), DB) %>%
  mutate(DIAG_CRS = "CRS")
dim(prevdf.taxname.CRS)
```

### Non-CRS
```{r}
# Compute prevalence of each feature, store as data.frame
prevdf.taxname.nonCRS = apply(X = otu_table(nonCRS.taxname),
               MARGIN = ifelse(taxa_are_rows(nonCRS.taxname), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
# Add taxonomy and total read counts to this data.frame
prevdf.taxname.nonCRS = data.frame(Prevalence = prevdf.taxname.nonCRS,
                    Abundance = taxa_sums(nonCRS.taxname),
                    phyloseq::tax_table(nonCRS.taxname)@.Data) %>%
  mutate(PercentAbundance = round(Abundance/sum(Abundance) * 100, 4)) %>%
  rownames_to_column("ASV") %>%
  left_join(nonCRS.taxname.prop.melt, by = "TaxName") %>%
  column_to_rownames("ASV") %>%
  select(Prevalence, Abundance, PercentAbundance, avgRelAbun, TaxName, everything()) %>%
  mutate(DIAG_CRS = "NonCRS")
```

## Filter to include just the top 20 taxa
```{r}
prevdf.taxname.CRS.top20genera <- filter(prevdf.taxname.CRS, Genus %in% top20generaNames)
```
```{r}
prevdf.taxname.nonCRS.top20genera <- filter(prevdf.taxname.nonCRS, Genus %in% top20generaNames)
```

## Bind the two tables together
```{r}
prevdf.taxname.top20genera <- bind_rows(prevdf.taxname.CRS.top20genera, prevdf.taxname.nonCRS.top20genera) %>%
  pivot_wider(id_cols = TaxName:Genus_Species2, names_from = DIAG_CRS, values_from = c("Prevalence","Abundance","PercentAbundance", "avgRelAbun")) %>%
  select(Kingdom, Phylum, Class, Order, Family, Genus, Species, TaxName, 
         Prevalence_CRS, Prevalence_NonCRS, Abundance_CRS, Abundance_NonCRS, 
         PercentAbundance_CRS, PercentAbundance_NonCRS, avgRelAbun_CRS, avgRelAbun_NonCRS) %>%
  arrange(Genus, TaxName)
```
Bind database info to table
```{r}
taxname.db <- data.frame(tax_table(data))
taxname.db <- select(taxname.db, TaxName, DB) %>%
  group_by(TaxName) %>%
  dplyr::summarise(DB=paste(DB, collapse = "/")) %>%
  ungroup() %>%
  mutate(DB = ifelse(grepl("Both", DB), "Both",
                     ifelse(grepl("NA", DB), "NA",
                            ifelse(grepl("SILVA/SILVA",DB), "SILVA", 
                                   ifelse(grepl("HOMD/SILVA",DB), "Both",DB)))))

prevdf.taxname.top20genera.db <- left_join(prevdf.taxname.top20genera, taxname.db, by = "TaxName")
```
```{r}
#write_csv(prevdf.taxname.top20genera.db,
#          "16S_output/top20PrevAbun.csv")
```


```{r}
library(gt)
library(gtsummary)
library(webshot)
```

```{r}
tab <- prevdf.taxname.top20genera.db %>%
  select(Genus, TaxName, Prevalence_CRS, Prevalence_NonCRS, avgRelAbun_CRS, avgRelAbun_NonCRS, DB) %>%
  gt(rowname_col = "TaxName", groupname_col = "Genus") %>%
  fmt_number(columns = vars(avgRelAbun_CRS, avgRelAbun_NonCRS), decimals = 2) %>%
  fmt_missing(columns = everything()) %>%
  summary_rows(groups = TRUE, 
               columns = vars(Prevalence_CRS, Prevalence_NonCRS, avgRelAbun_CRS, avgRelAbun_NonCRS),
               fns = list(Total = ~sum(., na.rm = TRUE)),
               use_seps = FALSE) %>%
  cols_align(align = "center", columns = vars(Prevalence_CRS, Prevalence_NonCRS, avgRelAbun_CRS, avgRelAbun_NonCRS, DB)) %>%
  tab_stubhead(label = md("**Taxonomic Feature**")) %>%
  tab_spanner(label = md("**Prevalence**"),
              columns = vars(Prevalence_CRS, Prevalence_NonCRS)) %>%
  tab_spanner(label = md("**Average % Abundance**"),
              columns = vars(avgRelAbun_CRS, avgRelAbun_NonCRS)) %>%
  cols_label(Prevalence_CRS = md("**CRS**"),
             Prevalence_NonCRS = md("**Non-CRS**"),
             avgRelAbun_CRS = md("**CRS**"),
             avgRelAbun_NonCRS = md("**Non-CRS**"),
             DB = md("**Database**")) %>%
  tab_style(style = cell_text(size = px(9), color = "black", font = "arial"),
    locations = cells_body(vars(Prevalence_CRS, Prevalence_NonCRS, avgRelAbun_CRS, avgRelAbun_NonCRS, DB))) %>%
  tab_style(style = cell_text(size = px(9), color = "black", font = "arial"),
    locations = cells_summary(groups = TRUE, rows = TRUE, columns = TRUE)) %>%
  tab_style(style = cell_text(size = px(9), color = "black", font = "arial", style = "italic", weight = "bold"),
    locations = cells_row_groups(groups = TRUE)) %>%
  tab_style(style = cell_text(size = px(9), color = "black", font = "arial", style = "italic"),
    locations = cells_stub(rows = TRUE)) %>%
  tab_options(
    table.font.size = 9,
    table.font.names = "arial",
    data_row.padding = px(2),
    summary_row.padding = px(2),
    row_group.padding = px(2),
    table.border.top.style = "none",
    table.border.bottom.style = "none",
    table_body.border.top.style = "none",
    table_body.border.bottom.style = "none",
  )

tab
#gtsave(tab, filename = "16S_output/Top20prevAbundTable.pdf")
```

## Export as a .CSV
```{r}
#write_csv(prevdf.taxname.top20genera, "16S_output/top20PrevAbun.csv")
```

# Relative Abundance Boxplots

## List of bacteria of interest
```{r}
Anaerobes <- c("Fusobacterium", "Prevotella", "Parvimonas", "Peptostreptococcus", "Cutibacterium", "Leptotrichia", "Anaerococcus","Peptoniphilus","Veillonella","Akkermansia","Finegoldia","Clostridium","Bacteroides","Porphyromonas","Actinomyces","Streptococcus","Selenomonas","Tannerella", "Dialister","Capnocytophaga", "Gemella","Eikenella", "Solobacterium", "Campylobacter","Atopobium", "Oribacterium","Treponema")
```

## Agglomerate data to the Genus level.
```{r}
anaerobes.prop <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(Genus %in% Anaerobes) %>%
  psmelt()
```

## Anaerobe Exploratory Boxplots by CRS Diagnosis
```{r fig.height=10, fig.width=16}
p <- ggboxplot(anaerobes.prop, 
              x = "DIAG_CRS", y = "Abundance",
              #add = "jitter",
              facet.by = "Genus",
              scales = "free"
              ) + 
  stat_compare_means(aes(group = DIAG_CRS), method = "wilcox.test") + 
  scale_y_log10()
p
```


```{r}
plot_abundance = function(physeq,title = "",
			     Facet = "Genus", Color = "DIAG_CRS"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(physeq, Phylum %in% c("Bacteroidetes","Fusobacteria"))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "DIAG_CRS",y = "Abundance",
                                 color = Color, fill = Color)) +
    geom_boxplot(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
                position = position_jitter(width = 0.3)) +
    theme_minimal() +
    facet_wrap(facets = Facet, scales = "free") + 
    scale_y_log10() +
    theme(legend.position="none",
          axis.text.x = element_text(angle = 90))
}
```

transform data
```{r}
CRSprop <- transform_sample_counts(CRS, function(x) x/sum(x) * 100)
NonCRSprop <- transform_sample_counts(nonCRS, function(x) x/sum(x) * 100)
```


```{r}
dataProp <- transform_sample_counts(data, function(x) x/sum(x)*100)
dataLog <- transform_sample_counts(data, function(x) log(1 + x))
```

```{r}
TopTaxPropPlotCRS <- plot_abundance(dataProp, Facet = "Genus", Color = "DIAG_CRS")
TopTaxPropPlotCRS
```

# Statistical analysis of Bacterial taxa with clinical parameters
```{r}
clinicalData <- data.frame(sample_data(data))
clinicalData
```

Which Phyla are assoicated with CRS and Non-CRS samples
```{r}
dataPhylumPropMeltFESS <- data %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  psmelt() %>%
  filter(Abundance > 0)
```
```{r}
ggboxplot(dataPhylumPropMeltFESS, x = as.character("DIAG_CRS"), y = "Abundance",
          facet.by = "Phylum",
          scales = "free",
          palette = "npg",      
          ) + 
  stat_compare_means(method = "wilcox.test", size = 3) + 
  scale_y_log10() +
  theme_pubr(base_size = 10)
```
Actinobacteria are significantly increased in abundance in Non-CRS samples. Is Actinobacteria abundance associated with any of the clinical factors?
## Actinobacteria
Narrow in on Actino
```{r}
Actino <- data %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Phylum == "Actinobacteria")
```
Get number of CRS and non CRS to report "n" in paper
```{r}
Actino %>%
  group_by(DIAG_CRS) %>%
  tally()
```
Actinobacteria abundance in CRS and non-CRS samples:
```{r fig.height=2.1, fig.width=2.5}
ActinoDIAGboxplot <- ggboxplot(Actino, 
                               x = "DIAG_CRS", 
                               y = "Abundance",
                               fill = "#DE2D26", 
                               add = "jitter", add.params = list(size=1, alpha = 0.6),
                               xlab = "", ylab = "Actinobacteria (%)") + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 1.4, label.y = 1.8, size = 6) + 
  scale_y_log10() +
  theme_pubr(base_size = 10, base_family = "Arial") +
  theme(legend.position  = "",
        plot.margin = margin(0.5,1,0,0, "cm")) +
  font("ylab", size = 11, color = "black")
ActinoDIAGboxplot
```

Actinobacteria Abundance and incidence of prior FESS surgery in CRS subjects
```{r}
ActinoFESS <- filter(Actino, !is.na(SINUS_FACTORS_PRIORFESS),
                     DIAG_CRS == "CRS") %>%
  mutate(SINUS_FACTORS_PRIORFESS = ifelse(SINUS_FACTORS_PRIORFESS == 1, "Yes",
                                          ifelse(SINUS_FACTORS_PRIORFESS == 0, "No", NA)))
```
Get number of CRS and non CRS to report "n" in paper
```{r}
ActinoFESS %>%
  group_by(SINUS_FACTORS_PRIORFESS) %>%
  tally()
```
Plot ActinoFESS SINUS_FACTORS_PRIORFESS
```{r fig.height=2.1, fig.width=2.5}
# New levels for axis order
ActinoFESS$SINUS_FACTORS_PRIORFESS = factor(ActinoFESS$SINUS_FACTORS_PRIORFESS, levels=c("Yes","No"))

ActinoFESSboxplot <- ggboxplot(ActinoFESS, 
                               x = as.character("SINUS_FACTORS_PRIORFESS"), 
                               y = "Abundance",fill = "#DE2D26", 
                               add = "jitter", add.params = list(size=1, alpha = 0.6),
                               xlab = "", ylab = "Actinobacteria (%)") + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x = 1.4, label.y = 1.8, size = 6) + 
  scale_y_log10() +
  theme_pubr(base_size = 10, base_family = "Arial") +
  theme(legend.position  = "",
        plot.margin = margin(0.2,1,0,0, "cm")) +
  font("ylab", size = 11, color = "black")
ActinoFESSboxplot
```
```{r fig.height=2.2, fig.width=5}
ActinoGrid <- plot_grid(ActinoDIAGboxplot, ActinoFESSboxplot, ncol = 2, align = "hv", rel_widths = c(1, 1))
ActinoGrid
```

```{r}
#ggsave(ActinoGrid,
#       filename = "16S_output/figures/ActinoBoxplotGrid.pdf",
#       device = cairo_pdf,
#       bg = "transparent",
#       height = 2.2,
#       width = 5
#       )
```
Actinobacteria abundance is significantly higher in CRS subjects with no history of previous FESS.  

Is Actinobacteria abundance correlated with SNOT20 scores?
```{r}
ActinoSNOT20Scatter <- ggscatter(Actino, x = "SNOT20_TOTAL", y = "Abundance",
          add = "reg.line",                         # Add regression line
          conf.int = TRUE,                          # Add confidence interval
          shape = "DIAG_CRS",
          palette = "npg",           
          ) +
  ggtitle("Actinobacteria Abundance and SNOT22 Scores") +
  labs(x = "SNOT20 Total", y = "Relative Abundance (%)") +
  scale_y_log10() +
  theme_pubr(base_size = 10) +
  stat_cor(method = "spearman")  # Add correlation coefficient
ActinoSNOT20Scatter
```

### Corynebacterium
Within Actinobacteria, in CRS subjects, is abundance of the genus Corynebacterium negatively associated with a history of previos FESS?
```{r}
Coryne <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus == "Corynebacterium")
```

```{r}
CoryneFESS <- filter(Coryne, !is.na(SINUS_FACTORS_PRIORFESS),
                     DIAG_CRS == "CRS")

CoryneFESSboxplot <- ggboxplot(CoryneFESS, x = as.character("SINUS_FACTORS_PRIORFESS"), y = "Abundance",
          #color = as.character("SINUS_FACTORS_PRIORFESS"),
          fill = as.character("SINUS_FACTORS_PRIORFESS"),
          add = "jitter",
          palette = "npg",  
          title = "Corynebacterium abundance \nin patients with prior FESS"
          ) + 
  stat_compare_means(method = "wilcox.test") + 
  scale_y_log10() +
  scale_x_discrete(labels = c("No","Yes")) +
  labs(x = "Prior FESS", y = "Relative Abundance (%)") +
  theme_pubr(base_size = 10) +
  theme(legend.position  = "")
CoryneFESSboxplot
```
The relationship is not significant.

Is Corynebacterium abundance correlated with SNOT20 scores?
```{r}
CoryneSNOT20Scatter <- ggscatter(Coryne, x = "SNOT20_TOTAL", y = "Abundance",
          add = "reg.line",                         # Add regression line
          conf.int = TRUE,                          # Add confidence interval
          shape = "DIAG_CRS",
          palette = "npg",           
          ) +
  ggtitle("Corynebacterium Abundance and SNOT22 Scores") +
  labs(x = "SNOT20 Total", y = "Relative Abundance (%)") +
  scale_y_log10() +
  theme_pubr(base_size = 10) +
  stat_cor(method = "spearman")  # Add correlation coefficient
CoryneSNOT20Scatter
```

### Rothia

```{r}
Rothia <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus == "Rothia")
```

### Cutibacterium
```{r}
Cuti <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus == "Cutibacterium")
```

```{r}
CutiDIAGboxplot <- ggboxplot(Cuti, x = "DIAG_CRS", y = "Abundance",
          #color = as.character("SINUS_FACTORS_PRIORFESS"),
          fill = "DIAG_CRS",
          add = "jitter",
          palette = "npg",  
          title = "Cutibacterium abundance and \nCRS diagnosis"
          ) + 
  stat_compare_means(method = "wilcox.test") + 
  scale_y_log10() +
  scale_x_discrete(labels = c("CRS","Non-CRS")) +
  labs(x = "", y = "Relative Abundance (%)") +
  theme(legend.position  = "") +
  theme_pubr(base_size = 10)
CutiDIAGboxplot
```

### Dolosigranulum
```{r}
Dolo <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus == "Dolosigranulum")
```

### "Anaerobes"

Could also add a level called anaerobe to the taxatable....
```{r}
AnaerobesDF <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus %in% Anaerobes)
```

```{r}
AnaerobesGroupSum <- AnaerobesDF %>%
  group_by(DIAG_CRS, SAMPLE_NAME) %>%
  dplyr::summarise(SumAnaerobes = sum(Abundance),
                   MeanAnaerobes = mean(Abundance))
```

```{r}
AnaerobeSumDIAGboxplot <- ggboxplot(AnaerobesGroupSum, x = "DIAG_CRS", y = "MeanAnaerobes",
          #color = as.character("SINUS_FACTORS_PRIORFESS"),
          fill = "DIAG_CRS",
          add = "jitter",
          palette = "npg",  
          title = "Anaerobe abundance and \nCRS diagnosis"
          ) + 
  stat_compare_means(method = "wilcox.test") + 
  scale_y_log10() +
  scale_x_discrete(labels = c("CRS","Non-CRS")) +
  labs(x = "", y = "Relative Abundance (%)") +
  theme(legend.position  = "") +
  theme_pubr(base_size = 10)
AnaerobeSumDIAGboxplot
```
### Streptococci

Could also add a level called anaerobe to the taxatable....
```{r}
Strep <- data %>%
  tax_glom("Genus") %>%
  transform_sample_counts(function(x) x/sum(x) *100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Genus %in% "Streptococcus")
```

```{r}
StrepDIAGboxplot <- ggboxplot(Strep, x = "DIAG_CRS", y = "Abundance",
          #color = as.character("SINUS_FACTORS_PRIORFESS"),
          fill = "DIAG_CRS",
          add = "jitter",
          palette = "npg",  
          title = "Streptococcus abundance \nand CRS diagnosis"
          ) + 
  stat_compare_means(method = "wilcox.test") + 
  scale_y_log10() +
  scale_x_discrete(labels = c("CRS","Non-CRS")) +
  labs(x = "", y = "Relative Abundance (%)") +
  theme(legend.position  = "") +
  theme_pubr(base_size = 10)
StrepDIAGboxplot
```
Strep is not significantly different.

Is Strep abundance correlated with SNOT20 scores?
```{r}
StrepSNOT20Scatter <- ggscatter(Strep, x = "SNOT20_TOTAL", y = "Abundance",
          add = "reg.line",                         # Add regression line
          conf.int = TRUE,                          # Add confidence interval
          shape = "DIAG_CRS",
          palette = "npg",           
          ) +
  ggtitle("Streptococcus Abundance and SNOT22 Scores") +
  labs(x = "SNOT20 Total", y = "Relative Abundance (%)") +
  scale_y_log10() +
  theme_pubr(base_size = 10) +
  stat_cor(method = "spearman")  # Add correlation coefficient
StrepSNOT20Scatter
```

### Actinobacteria correlation with Bacteroidetes and Fusobacteria
Now I want to see if Bacteroidetes is negatively correlated with Actinobacteria. 
Create a dataset with just Actinobacteria and Bacteroidetes
```{r}
ActinoBactFuso_df <- data  %>%
  tax_glom("Phylum") %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  psmelt() %>%
  filter(Abundance > 0,
         Phylum %in% c("Actinobacteria","Bacteroidetes","Fusobacteria")) %>%
  select(DIAG_CRS, SAMPLE_NAME, Phylum, Abundance) %>%
  pivot_wider(names_from = Phylum, values_from = Abundance) %>%
  mutate(logActino = log(Actinobacteria),
         logBact = log(Bacteroidetes),
         logFuso = log(Fusobacteria))
```
Now, make a scatterplot of Actino and Bacteroidetes in each sample
```{r}
ActinoBactScatter <- ggscatter(ActinoBactFuso_df, x = "Bacteroidetes", y = "Actinobacteria",
          add = "reg.line",                         # Add regression line
          conf.int = TRUE,                          # Add confidence interval
          #color = "DIAG_CRS",
          palette = "npg",           # Color by groups "cyl"
          ) +
  scale_y_log10() +
  scale_x_log10() +
  stat_cor(method = "spearman")  
```

Now, make a scatterplot of Actino and Bacteroidetes in each sample
```{r}
ActinoFusoScatter <- ggscatter(ActinoBactFuso_df, x = "Fusobacteria", y = "Actinobacteria",
          add = "reg.line", 
          conf.int = TRUE, 
          palette = "npg",           
          ) +
  scale_y_log10() +
  scale_x_log10() +
  stat_cor(method = "spearman")  
```

```{r}
ActinoBactFuso_grid <- plot_grid(ActinoBactScatter, ActinoFusoScatter, ncol = 2, align = 'hv')
ActinoBactFuso_grid
```

```{r}
#ggsave(ActinoBactFuso_grid,
#       filename = "16S_output/figures/ActinoBactFusoScatterGrid.pdf",
#       device = "pdf",
#       width = 6, height = 3,
#      useDingbats = FALSE)
```

