---
title: "FPLC Unicorn File Analysis"
Author: "Sarah Lucas"
---
# Steps to get from Unicorn formatted files to R plots

## Step 1. Download PyCorn. Most computers already have python (at least 2.7, but probably 2.7/3.x) pre-installed so copy and past the following code into a Terminal:

pip install pycorn

## Step 2. Export each Unicorn file as a .zip file and transfer them to your computer. Save all of them in a single folder. Enter that folder in your terminal. There are options for exporting files directly to .csv files, but I haven't been able to make it work yet. the pycorn-bin.py does havea csv flag instead of xlsx, but I haven't been able to make that work. So, Copy and past the following into your terminal with *zip ("all .zip files") as the 'input.res' option:

* Extract data to xlsx-file:

pycorn-bin.py -e xlsx input.res <-- DUMMY EXAMPLE
pycorn-bin.py -e xlsx *zip <-- COPY AND PASTE

Then, you can batch convert the xlsx files 

* download the python package xlsx2csv:

pip install xlsx2csv

* make a script named batch_xlsx2csv.sh containing the text below and save it somewhere in your PATH (don't forget to de-comment the lines):

for i in *.xlsx;
 do
  filename=$(basename "$i" .xlsx);
  outext=".csv"
  xlsx2csv $i $filename$outext
done

* make the script executable:

chmod +x batch_xlsx2csv.sh to 

Go back to the folder with your new xlsx files and run the line:
batch_xlsx2csv.sh

## Step 3. Load the libararies below.
```{r}
library(DescTools)
library(tidyverse)
library(ggpubr)
```
Aesthetics
```{r}
cbPal_MMM <- c("black","pink","#E69F00","#56B4E9","#009E73","#F0E442", "#0072B2","#75437C","#B52F25") 
```

## Step 4a. Function for importing .csv files
```{r}
read_chr_data <- function (x) {
  read_csv(x, col_types = 
             cols(
               UV = col_double(),
               X18 = col_double()
             )) %>%
  data.frame() %>%
  select(ElutionVolume = UV, Absorbance280 = X18) %>%
  slice(49:39797)
}
```
## Step 4b. Import csv files. Customize to file location and data object naming.
```{r}
chr_data_MMM <- read_chr_data("../data/enrichment/fplc/MMM_blank.csv")
chr_data_164 <- read_chr_data("../data/enrichment/fplc/Super164_blank.csv")
chr_data_222 <- read_chr_data("../data/enrichment/fplc/Super222_blank.csv")
chr_data_230 <- read_chr_data("../data/enrichment/fplc/Super230_blank.csv")
chr_data_234 <- read_chr_data("../data/enrichment/fplc/Super234_blank.csv")
chr_data_235 <- read_chr_data("../data/enrichment/fplc/Super235_blank.csv")
chr_data_243 <- read_chr_data("../data/enrichment/fplc/Super243_blank.csv")
chr_data_260 <- read_chr_data("../data/enrichment/fplc/Super260_blank.csv")
chr_data_262 <- read_chr_data("../data/enrichment/fplc/Super262_blank.csv")
```

## Step 5. Customizable. Add/delete data objects from the plot. Change colors, aesthetics, etc.
```{r}
chr_data_MMM_first_peak <- chr_data_MMM %>% filter(between(ElutionVolume, 7, 10))
chr_data_164_first_peak <- chr_data_164 %>% filter(between(ElutionVolume, 7, 10))
chr_data_222_first_peak <- chr_data_222 %>% filter(between(ElutionVolume, 7, 10))
chr_data_230_first_peak <- chr_data_230 %>% filter(between(ElutionVolume, 7, 10))
chr_data_234_first_peak <- chr_data_234 %>% filter(between(ElutionVolume, 7, 10))
chr_data_235_first_peak <- chr_data_235 %>% filter(between(ElutionVolume, 7, 10))
chr_data_243_first_peak <- chr_data_243 %>% filter(between(ElutionVolume, 7, 10))
chr_data_260_first_peak <- chr_data_260 %>% filter(between(ElutionVolume, 7, 10))
chr_data_262_first_peak <- chr_data_262 %>% filter(between(ElutionVolume, 7, 10))
```
## Plot
```{r fig.height=3, fig.width=3}
chromatogram_firstpeak <- ggplot(chr_data_MMM_first_peak, aes(x=as.numeric(ElutionVolume), y = as.numeric(Absorbance280))) +
  geom_line() +
  geom_step(data = chr_data_164_first_peak, color = "pink") +
  geom_step(data = chr_data_222_first_peak, color = "#E69F00") +
  geom_step(data = chr_data_230_first_peak, color = "#56B4E9") +
  geom_step(data = chr_data_234_first_peak, color = "#009E73") +
  geom_step(data = chr_data_235_first_peak, color = "#F0E442") +
  geom_step(data = chr_data_243_first_peak, color = "#0072B2") +
  geom_step(data = chr_data_260_first_peak, color = "#75437C") +
  geom_step(data = chr_data_262_first_peak, color = "#B52F25") +
  geom_vline(xintercept = 7.2, alpha = 0.5, linetype = 2) +
  geom_vline(xintercept = 8.2, alpha = 0.5, linetype = 2) +
  scale_x_continuous(name = "Elution Volume (mL)") +
  ylab("Absorbance (280nm)") +
  theme_pubr(base_size = 10)

chromatogram_firstpeak
```
```{r}
ggsave(plot = chromatogram_firstpeak, 
       file = "enrichment_output/fplc_firstpeak.pdf", 
       device = "pdf", 
       width = 2.5, height = 2.5)
```

```{r}
sessionInfo()
```

