---
title: "S. aureus LAC aerobic/anaerobic growth on MMM"
author: "Sarah Lucas"
---

# Setup
## Packages
```{r}
library(tidyverse)
library(lubridate)
library(cowplot)
library(scales)
library(hms)
library(ggpubr)
library(extrafont)
```

## Plotting
```{r}
cbPal <- c("pink",  # 164
           "#E69F00", # 222
           "#56B4E9", # 230
           "#009E73", # 234
           "#F0E442", # 235
           "#0072B2", # 243
           "#D55E00", # 260
           "red") #262
```


# Aerobic Growth
## Experiemntal Parameters to fill in: Media in plate
```{r}
Medium.A <- "MMM"
Medium.B <- "MMM+Lactate(0.5%)"
Medium.C <- "MMM+CAA(0.5%)"
Medium.D <- "MMM+Glucose(0.25%)"
Medium.E <- "MMM+CAA+Lactate"
Medium.F <- "MMM+CAA+Glucose"
Medium.G <- "MMM+Glucose+NH3"
Medium.H <- "MMM+CAA+Glucose+NH3"
```
## Experiemntal Parameters to fill in: Samples in wells
```{r}
Bio1 <- "LAC1"
Bio2 <- "LAC2"
Bio3 <- "LAC3"
Bio4 <- "Medium_Control"
```
## Import data
```{r}
start_time <- as.hms(00:00:00)

AerobicData <- read_csv("../data/enrichment/USA300_growth/Growth_Curve_MMM_Add_12MAR2020.csv", skip = 46) %>%
  select(-contains("600")) %>%
  filter(!is.na(Time)) %>%
  gather(key = Well, value = OD600, A1:H12) %>%
  separate(Well, into = c("rowx","Column"), sep = "[[:alpha:]]", remove = FALSE) %>%
  separate(Well, into = c("Row","columnx"), sep = "[[:digit:]]", remove = FALSE) %>%
  select(Time, Well, Row, Column, OD600) %>%
  mutate(Medium = ifelse(Row == "A", Medium.A,
                         ifelse(Row == "B", Medium.B, 
                                ifelse(Row == "C", Medium.C, 
                                       ifelse(Row == "D", Medium.D,
                                              ifelse(Row == "E", Medium.E, 
                                                     ifelse(Row == "F", Medium.F, 
                                                            ifelse(Row == "G", Medium.G, Medium.H)))))))) %>%
  mutate(BioRep = ifelse(Column %in% 1:3, Bio1,
                         ifelse(Column %in% 4:6, Bio2,
                                ifelse(Column %in% 7:9, Bio3,
                                       Bio4)))) %>%
  mutate(difftime = as.numeric(abs(difftime(start_time, Time, units = "hours"))))
AerobicData$Column <- as.numeric(AerobicData$Column)
```

## Growth Curves by well
```{r}
GC_plate_plot <- ggplot(AerobicData, aes(x=difftime, y=OD600, color = Medium)) + 
  facet_grid(Row ~ Column) +
  theme_classic() + 
  geom_line(size = 1) +
  scale_color_manual(values = c(cbPal)) +
  scale_y_log10() +
  annotation_logticks(base = 10, 
                      sides = "l"
  ) +
  labs(x = "", y = "") +
  theme(legend.position = "right", 
        axis.text.x = element_text(color = "black", size = 6, face = "bold", angle = 90),
        axis.text.y = element_text(color = "black", size = 6, face = "bold")
  )
GC_plate_plot
```
## Normalize
Subtract data blanks at each time point for each medium.
```{r}
data_samples <- AerobicData %>%
  filter(BioRep != "Medium_Control")
  
data_blanks <- AerobicData %>%
  filter(BioRep == "Medium_Control") %>%
  group_by(Medium, Time) %>%
  # Get the average blank value for each supernatant at each timepoint
  summarize(blank_mean_OD600 = mean(OD600))

#Combine data back together
data_join <- left_join(data_samples, data_blanks)
data_join_blank <- data_join %>%
  mutate(blanked_OD600 = OD600 - blank_mean_OD600)
```
## Summarize
Get the mean and standard deviation for each supernatant at each timepoint for plotting
```{r}
data_join_summarized <- data_join_blank %>%
  group_by(Medium, Time, difftime) %>%
  summarize(mean_blanked_OD600 = mean(blanked_OD600), sd_blanked_OD600 = sd(blanked_OD600)) %>%
  ungroup()
data_join_summarized$Time <- hms::as.hms(data_join_summarized$Time)
```
## Plot Summary Growth Curves

### Filter for plotting points and error bars at every hour instead of every half hour.
```{r}
data_join_summarized_error <- filter(data_join_summarized, minute(Time) == 29)
```
### Remove the NH3 containing conditions
```{r}
NitrateMedia <- c("MMM+Glucose+NH3", "MMM+CAA+Glucose+NH3")
data_join_summarized_noNH3 <- filter(data_join_summarized, !Medium %in% NitrateMedia)
data_join_summarized_error_noNH3 <- filter(data_join_summarized_error, !Medium %in% NitrateMedia)
```

### Plot
Growth curves are plotted at half hour intervals, while points and error bars are plotted at hourly intervals to avoid overplotting and enable readability of the graph.
```{r}
# plotting points and error bars for when they overlap
pd <- position_dodge(0.1)
# Plotting breaks
breaks = c("MMM","MMM+CAA(0.5%)","MMM+Lactate(0.5%)","MMM+CAA+Lactate","MMM+Glucose(0.25%)","MMM+CAA+Glucose")
labels = c("MMM","MMMC","MMML","MMMLC","MMMG","MMMGC")
#majorY <- 10^(-3:0)
#minorY <- rep(1:9, 21)*(10^rep(-3:-1, each=9))

# Plot
GC_plot_sd <- ggplot(data_join_summarized_noNH3) +
  geom_line(aes(x=difftime, y=mean_blanked_OD600, 
                color = Medium), 
            size = .5) +
  geom_errorbar(data = data_join_summarized_error_noNH3, 
                mapping = aes(x = difftime, 
                              ymin = mean_blanked_OD600-sd_blanked_OD600, 
                              ymax = mean_blanked_OD600+sd_blanked_OD600), 
                position = pd, size = .5, alpha = .6, width = 0.2) +
  geom_point(data = data_join_summarized_error_noNH3,
             size = 2,
             mapping = aes(x = difftime, 
                           y = mean_blanked_OD600,
                           shape = Medium,
                           fill = Medium)) +
  scale_shape_manual(breaks = breaks,
                     labels = labels,
                     values = c(21,21,22,22,24,24)
  ) +
  scale_color_manual(breaks = breaks,
                     labels = labels,
                     values = c("black","black","black","black","black","black")) +
  scale_fill_manual(breaks = breaks,
                    labels = labels,
                    values = c("black","white","black","white","black","white")) +
  scale_y_log10(limits =c(.01, 0.6), 
                breaks = c(.01, .1, .6)
                #breaks = majorY,
                #minor_breaks = minorY
                ) +
  annotation_logticks(base = 10, sides = "l",
                      color = "black"
  ) +
  scale_x_continuous(limits = c(0,18), 
                     breaks = c(0,3,6,9,12,15,18,20)) +
  labs(title = "Aerobic", 
       x = "Time (hrs)", 
       y = expression(paste("O.D."[600][nm]))) +
  theme_pubr(base_size = 10, base_family = "Arial") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, vjust = 1, size = 10),
        plot.margin = unit(c(0.1,1,.6,0), "cm"),
        axis.ticks = element_line(color = "black"),
        axis.line = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black"),
        legend.position = c(.8, .25),
        legend.key.size = unit(0.35,"cm"),
        legend.text = element_text(size = 8),
        legend.background = element_blank(),
        legend.title = element_blank())

GC_plot_sd
```


# Anaerobic Growth
## Experiemntal Parameters to fill in: Media in plate
```{r}
AnaerobicMedium.A <- "MMM"
AnaerobicMedium.B <- "MMM+CAA(0.5%)"
AnaerobicMedium.C <- "MMM+Lactate(0.5%)"
AnaerobicMedium.D <- "MMM+Glucose(0.25%)"
AnaerobicMedium.E <- "MMM+CAA+Lactate"
AnaerobicMedium.F <- "MMM+CAA+Glucose"
```
## Experimental Parameters to fill in: Samples in wells
```{r}
Bio1 <- "LAC1"
Bio2 <- "LAC2"
Bio3 <- "LAC3"
Bio4 <- "Medium_Control"
```

## Import Anaerobic plate data
```{r}
anaerobicData <- read_csv(file = "../data/enrichment/USA300_growth/Anaerobic_Growth_LAC_APR2020.csv") %>%
  select(Row = X1, everything()) %>%
  pivot_longer(-Row, names_to = "Column", values_to = "OD600") %>%
  mutate(BioRep = ifelse(Column %in% 1:3, Bio1,
                         ifelse(Column %in% 4:6, Bio2,
                                ifelse(Column %in% 7:9, Bio3, Bio4))),
         Medium = ifelse(Row == "A", AnaerobicMedium.A,
                         ifelse(Row == "B", AnaerobicMedium.B, 
                                ifelse(Row == "C", AnaerobicMedium.C, 
                                       ifelse(Row == "D", AnaerobicMedium.D,
                                              ifelse(Row == "E", AnaerobicMedium.E, AnaerobicMedium.F))))),
         GrowthCondition = "Anaerobic")
```
### Plate Heatmap 
Make a heatmap of the plate to identify any outliers
```{r}
anaerobicPlateHeat <- ggplot(anaerobicData, mapping = aes(x = Column, y = Row)) +
  geom_tile(aes(fill = OD600), color = "whitesmoke") +
  geom_text(aes(label = OD600), color = "white") +
  scale_x_discrete(limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) +
  scale_y_discrete(limits = c("F","E","D","C","B","A")) +
  theme_pubclean(base_size = 14) +
  theme(legend.position = "right")
anaerobicPlateHeat
```
## Import Anaerobic + Nitrate (3mM) plate data
```{r}
anaerobicNitrateData <- read_csv(file = "../data/enrichment/USA300_growth/Anaerobic_Nitrate_Growth_LAC_APR2020.csv") %>%
  select(Row = X1, everything()) %>%
  pivot_longer(-Row, names_to = "Column", values_to = "OD600") %>%
  mutate(BioRep = ifelse(Column %in% 1:3, Bio1,
                         ifelse(Column %in% 4:6, Bio2,
                                ifelse(Column %in% 7:9, Bio3, Bio4))),
         Medium = ifelse(Row == "A", AnaerobicMedium.A,
                         ifelse(Row == "B", AnaerobicMedium.B, 
                                ifelse(Row == "C", AnaerobicMedium.C, 
                                       ifelse(Row == "D", AnaerobicMedium.D,
                                              ifelse(Row == "E", AnaerobicMedium.E, AnaerobicMedium.F))))),
         GrowthCondition = "Nitrate3mM")
anaerobicNitrateData
```
### Plate Heatmap 
Make a heatmap of the plate to identify any outliers
```{r}
nitratePlateHeat <- ggplot(anaerobicNitrateData, mapping = aes(x = Column, y = Row)) +
  geom_tile(aes(fill = OD600), color = "whitesmoke") +
  geom_text(aes(label = OD600), color = "white") +
  scale_x_discrete(limits = c("1","2","3","4","5","6","7","8","9","10","11","12")) +
  scale_y_discrete(limits = c("F","E","D","C","B","A")) +
  theme_pubclean(base_size = 14) +
  theme(legend.position = "right")
nitratePlateHeat
```

## Remove Outliers
Remove well A7 from analysis.
```{r}
anaerobicNitrateData <- filter(anaerobicNitrateData, !Row == "A" | !Column == "7")
```

## Bind the two datasets together:
```{r}
anaerobicDataBind <- bind_rows(anaerobicData, anaerobicNitrateData)
```


## Normalize
Subtract data blanks at each time point for each medium. Calculate average OD600 readings. Tidy data for plotting.
```{r}
anaerobic_samples <- anaerobicDataBind %>%
  filter(BioRep != "Medium_Control")

anaerobic_blanks <- anaerobicDataBind %>%
  filter(BioRep == "Medium_Control") %>%
  group_by(Medium, GrowthCondition) %>%
  summarize(blank_mean_OD600 = mean(OD600))

#Combine data back together, with a new column for averaged blanks
anaerobic_data_join <- left_join(anaerobic_samples, anaerobic_blanks, by = c("Medium","GrowthCondition"))

# For boxplots - blanked datapoints
anaerobic_data_blank <- anaerobic_data_join %>%
  mutate(blanked_OD600 = OD600 - blank_mean_OD600) %>%
  mutate(Date = "April2020") %>%
  select(Date, GrowthCondition, Medium, BioRep, OD600, blank_mean_OD600, blanked_OD600)
```

## Summarize
```{r}
# For barplots - summarized data for each medium/condition
anaerobic_data_summarize <- anaerobic_data_blank %>%
  group_by(Medium, ) %>%
  summarize(sd_blanked_OD600 = sd(blanked_OD600), blanked_OD600 = mean(blanked_OD600)) %>%
  ungroup()
```

## Statistical Testing
Compare each of the conditions to the MMM control in the Anaerobic and Anaerobic+Nitrate groups
```{r}
# Test difference in growth compared to MMM control

ttest_compare_group <- compare_means(blanked_OD600 ~ Medium,
                                     data = anaerobic_data_blank,
                                     ref.group = "MMM",
                                     method = "t.test",
                                     group.by = "GrowthCondition")
ttest_compare_group
```
## Plotting Summary Growth Data
```{r}
anaerobic_pubbarplot <- ggbarplot(anaerobic_data_blank, 
                        x = "Medium", y = "blanked_OD600",
                        add = c("mean_sd"),
                        facet.by = "GrowthCondition",
                        panel.labs = list(
                          GrowthCondition = c("Anaerobic", "Anaerobic + Nitrate (3mM)")),
                        scales = "free",
                        fill = "lightgrey",
                        ggtheme = theme_pubr(base_size = 10)) +
  scale_x_discrete(limits = c("MMM",
                              "MMM+CAA(0.5%)",
                              "MMM+Lactate(0.5%)",
                              "MMM+CAA+Lactate",
                              "MMM+Glucose(0.25%)",
                              "MMM+CAA+Glucose"),
                   labels = c("MMM",
                              "MMMC",
                              "MMML",
                              "MMMLC",
                              "MMMG",
                              "MMMGC")) +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0, .03, .06, .09, .12, .15, .18)) +
                     #breaks = c(0,.02,.04,.06, .08, .1, .12, .14, .16, .18, .2)) +
  theme_pubr(base_size = 14) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = expression(paste("O.D."[600][nm]))) +
  # add adjusted p-value stats to plot
  stat_pvalue_manual(ttest_compare_group, label = "p.signif", y.position = .2, remove.bracket = TRUE, color = "black")

anaerobic_pubbarplot 
```

I want to average these values with the previous anaerobic growth data from my August2019 experiment. The only difference between experiments is MMM batch.

## Import Anaerobic data from August 2019

```{r}
anaerobicAug2019 <- read_tsv(file = "../data/enrichment/USA300_growth/24Hr_Anaerobic_Growth_MMM-CAA-Lactate_30AUG2019.txt")
anaerobicAug2019
```

## Normalize
Subtract data blanks at each time point for each medium. Calculate average OD600 readings. Tidy data for plotting.
```{r}
anaerobicAug2019_samples <- anaerobicAug2019 %>%
  filter(BioRep != "Blank")

anaerobicAug2019_blanks <- anaerobicAug2019 %>%
  filter(BioRep == "Blank") %>%
  group_by(Medium, GrowthCondition) %>%
  summarize(blank_mean_OD600 = mean(OD600))

#Combine data back together, with a new column for averaged blanks
anaerobicAug2019_data_join <- left_join(anaerobicAug2019_samples, anaerobicAug2019_blanks)

# For boxplots - blanked datapoints
anaerobicAug2019_data_blank <- anaerobicAug2019_data_join %>%
  mutate(blanked_OD600 = OD600 - blank_mean_OD600) %>%
  mutate(Date = "Aug2020") %>%
  select(Date, GrowthCondition, Medium, BioRep, OD600, blank_mean_OD600, blanked_OD600)
```

Bind data together
```{r}
anaerobicBind <- bind_rows(anaerobic_data_blank, anaerobicAug2019_data_blank)
```

```{r}
# fig.height = 3, fig.width = 4, fig.alight="center"
anaerobicBind_pubbarplot <- ggbarplot(anaerobicBind, 
                        x = "Medium", y = "blanked_OD600",
                        add = c("mean_sd"),
                        facet.by = c("GrowthCondition", "Date"),
                        scales = "free",
                        fill = "lightgrey",
                        ggtheme = theme_pubr(base_size = 10)) +
  scale_x_discrete(limits = c("MMM",
                              "MMM+CAA(0.5%)",
                              "MMM+Lactate(0.5%)",
                              "MMM+CAA+Lactate",
                              "MMM+Glucose(0.25%)",
                              "MMM+CAA+Glucose"),
                   labels = c("MMM",
                              "MMMC",
                              "MMML",
                              "MMMLC",
                              "MMMG",
                              "MMMGC")) +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0,.02,.04,.06, .08, .1, .12,.14,.16,.18,2)) +
  theme_minimal(base_size = 20) +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = expression(paste("O.D."[600][nm]))) 
  
  # add adjusted p-value stats to plot
  #stat_pvalue_manual(anaerobic_ttest_compare, label = "p.signif", y.position = .15, remove.bracket = TRUE, color = "black")

anaerobicBind_pubbarplot 
```
Ok, these runs look really comparable so I feel OK averaging the numbers and adding the glucose conditions.

```{r}
# Test difference in growth compared to MMM control

ttest_compare_Bindgroup <- compare_means(blanked_OD600 ~ Medium,
                                     data = anaerobicBind,
                                     ref.group = "MMM",
                                     method = "t.test",
                                     group.by = "GrowthCondition")
ttest_compare_Bindgroup 
```

```{r}
# fig.height = 3, fig.width = 4, fig.alight="center"
anaerobicBind_pubbarplot <- ggbarplot(anaerobicBind, 
                        x = "Medium", y = "blanked_OD600",
                        add = c("mean_sd"),
                        facet.by = "GrowthCondition",
                        panel.labs = list(
                          GrowthCondition = c("Anaerobic", "Anaerobic + Nitrate (3mM)")),
                        scales = "free",
                        fill = "lightgrey",
                        ggtheme = theme_pubr(base_size = 10)) +
  scale_x_discrete(limits = c("MMM",
                              "MMM+CAA(0.5%)",
                              "MMM+Lactate(0.5%)",
                              "MMM+CAA+Lactate",
                              "MMM+Glucose(0.25%)",
                              "MMM+CAA+Glucose"),
                   labels = c("MMM",
                              "MMMC",
                              "MMML",
                              "MMMLC",
                              "MMMG",
                              "MMMGC")) +
  scale_y_continuous(limits = c(0,.2), 
                     breaks = c(0,.04,.08,.12,.16,.2)) +
  theme_pubr(base_size = 10, base_family = "Arial") +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.ticks = element_line(color = "black"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_rect(color = "white", fill = "white"),
        strip.text = element_text(face = "bold", size = 10, vjust = 2),
        plot.margin = unit(c(0,0,.35,0), "cm"),) +
  labs(y = expression(paste("O.D."[600][nm]))) +
  
  # add adjusted p-value stats to plot
  stat_pvalue_manual(ttest_compare_Bindgroup, 
                     label = "p.signif", 
                     y.position = c(.03,.035,.09,.04,.12,.045,.06,.12,.065,.2),
                     remove.bracket = TRUE, 
                     color = "black")

anaerobicBind_pubbarplot 
```

# Final Plot
```{r fig.height = 3, fig.width=8}
Figure4 <- plot_grid(GC_plot_sd, anaerobicBind_pubbarplot, 
                     align = "hv", 
                     labels = "auto", 
                     rel_widths = c(0.75, 1)
                     )
Figure4
```

```{r}
ggsave(Figure4, filename = "enrichment_output/Figure4.pdf",
       device = cairo_pdf,
       dpi = 300,
       #useDingbats = FALSE,
       width = 8,
       height = 3)
```
```{r}
ggsave(Figure4, filename = "enrichment_output/Figure4.png",
       device = "png",
       #useDingbats = FALSE,
       width = 8,
       height = 3)
```

